---
title: "📊DataPrepHIS Report"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: true
    number_sections: true
    theme: united
    code_folding: hide
    df_print: paged
    self_contained: true
    css: |
      .main-container {
        max-width: 1200px;
        margin-left: auto;
        margin-right: auto;
      }
      h1, h2, h3, h4 {
        margin-top: 24px;
        margin-bottom: 16px;
        font-weight: 600;
        line-height: 1.25;
      }
      ul, ol {
        padding-left: 2em;
        margin-bottom: 16px;
      }
      li {
        margin-bottom: 8px;
        display: list-item;
      }
      li ul, li ol {
        margin-top: 8px;
        margin-left: 1em;
      }
      ul li {
        list-style-type: disc;
      }
      ul li li {
        list-style-type: circle;
      }
      ul li li li {
        list-style-type: square;
      }
      ol li {
        list-style-type: decimal;
      }
      ol li li {
        list-style-type: lower-alpha;
      }
      ol li li li {
        list-style-type: lower-roman;
      }
      p {
        margin-bottom: 16px;
      }
      pre {
        margin-bottom: 16px;
      }
      .table-container {
        overflow-x: auto;
        margin-bottom: 20px;
      }
params:
  data: NULL
  original_filename: NULL
  preprocessing_summary: NULL
  numeric_vars: NULL
  categorical_vars: NULL
  visualizations: NULL
  include_plots: TRUE
  include_stats: TRUE
  include_preprocessing: TRUE
  include_correlation: TRUE
  include_quality_report: TRUE
  include_recommendations: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
library(dlookr)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(corrplot)
library(GGally)
library(moments)
library(htmltools)

# Helper function to safely handle errors in numeric calculations
safe_numeric <- function(expr, default = NA) {
  tryCatch(
    expr,
    error = function(e) default,
    warning = function(w) default
  )
}
```

# Data Pre-Processing and Visualization Report 

This report provides a detailed analysis of the data preprocessing steps and visualizations for the tool 📊DataPrepHIS.

## Dataset Overview

### Dataset Information

```{r results="asis"}
# Display dataset name
if(!is.null(params$original_filename)) {
  cat("<p><strong>Dataset Name:</strong> ", params$original_filename, "</p>")
} else {
  cat("<p><strong>Dataset Name:</strong> Not specified</p>")
}

# Add analysis date
cat("<p><strong>Analysis Date:</strong> ", format(Sys.Date(), "%B %d, %Y"), "</p>")

```

### Data Preview

```{r}
# Basic summary of dataset
if(!is.null(params$data) && nrow(params$data) > 0) {
  # Format data for display - handle potential formatting issues
  display_data <- params$data
  # Convert very large or very small numbers to a readable format
  display_data[] <- lapply(display_data, function(x) {
    if(is.numeric(x)) {
      # Format large numbers with commas and limit decimals
      ifelse(abs(x) > 1000000 | abs(x) < 0.001, 
             sprintf("%.2e", x),  # Scientific notation for very large/small
             x)  # Keep original for normal range
    } else {
      x  # Return non-numeric as is
    }
  })
  
  kable(head(display_data, 10), 
       caption = "First 10 rows of dataset") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                 full_width = TRUE) %>%
    scroll_box(width = "100%", height = "300px")
} else {
  cat("<p>No data available to display.</p>")
}
```

### Dataset Structure

```{r results="asis"}
# Display data structure
if(params$include_stats && !is.null(params$data) && ncol(params$data) > 0) {
  # Dataset dimensions
  cat("<p><strong>Dataset Dimensions:</strong> ", format(nrow(params$data), big.mark=","), " rows × ", ncol(params$data), " columns</p>\n\n")
}
```

## Data Preprocessing

### Preprocessing Operations

```{r}
# Display preprocessing summary table
if(params$include_preprocessing && !is.null(params$preprocessing_summary) && nrow(params$preprocessing_summary) > 0) {
  kable(params$preprocessing_summary, 
       caption = "Applied Preprocessing Operations") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
} else {
  cat("No preprocessing operations were recorded.")
}
```

### Missing Values Analysis

```{r}
# Display missing values analysis
if(params$include_stats && !is.null(params$data)) {
  # More robust check for missing values including empty strings and "NA" text
  any_missing <- sum(sapply(params$data, function(x) sum(is.na(x) | (is.character(x) & x == "") | (is.character(x) & x == "NA")))) > 0
  
  if(any_missing) {
    tryCatch({
      params$data %>% 
        plot_na_pareto(main = "Missing Values Distribution")
      
      # Add list of columns with missing values in the requested format
      missing_cols <- sapply(params$data, function(x) {
        if(is.factor(x) || is.character(x)) {
          sum(is.na(x) | x == "" | x == "NA", na.rm = TRUE)
        } else {
          sum(is.na(x), na.rm = TRUE)
        }
      })
      missing_cols <- missing_cols[missing_cols > 0]
      
      if(length(missing_cols) > 0) {
        cat("\n\n#### Columns with Missing Values\n\n")
        
        for(col_name in names(missing_cols)) {
          missing_count <- missing_cols[col_name]
          missing_pct <- round(100 * missing_count / nrow(params$data), 2)
          cat(col_name, " : ", missing_count, " values ( ", missing_pct, "% )\n\n", sep="")
        }
      }
    }, error = function(e) {
      cat("**Missing value analysis:** There are", sum(is.na(params$data)), "missing values in the dataset. Error details:", e$message)
    })
  } else {
    cat("**No missing values** found in the dataset.")
  }
}
```

## Data Exploration

### Summary Statistics

```{r}
if(params$include_stats && !is.null(params$data) && ncol(params$data) > 0) {
  # Create a summary table matching the Data Load tab format
  summary_df <- data.frame(
    #Column = names(params$data),
    Class = sapply(params$data, function(x) { if (is.numeric(x)) "Numerical" else "Categorical" }),
    Missing = sapply(params$data, function(x) {
      # Handle different data types for missing value detection
      if (is.factor(x) || is.character(x)) {
        sum(is.na(x) | x == "" | x == "NA", na.rm = TRUE)
      } else {
        sum(is.na(x), na.rm = TRUE)
      }
    }),
    Missing_Percent = sapply(params$data, function(x) {
      missing_count <- if (is.factor(x) || is.character(x)) {
        sum(is.na(x) | x == "" | x == "NA", na.rm = TRUE)
      } else {
        sum(is.na(x), na.rm = TRUE)
      }
      sprintf("%.1f%%", (missing_count / length(x)) * 100)
    }),
    Min = sapply(params$data, function(x) { if (is.numeric(x)) min(x, na.rm = TRUE) else "-" }),
    Median = sapply(params$data, function(x) { if (is.numeric(x)) median(x, na.rm = TRUE) else "-" }),
    Max = sapply(params$data, function(x) { if (is.numeric(x)) max(x, na.rm = TRUE) else "-" }),
    Mean = sapply(params$data, function(x) { if (is.numeric(x)) sprintf("%.3f", mean(x, na.rm = TRUE)) else "-" }),
    SD = sapply(params$data, function(x) { if (is.numeric(x)) sprintf("%.3f", sd(x, na.rm = TRUE)) else "-" }),
    Variance = sapply(params$data, function(x) { if (is.numeric(x)) sprintf("%.3f", var(x, na.rm = TRUE)) else "-" }),
    Unique_Values = sapply(params$data, function(x) { if (!is.numeric(x)) length(unique(x)) else "-" })
  )

  # Replace NAs with dash for display
  summary_df[] <- lapply(summary_df, function(x) ifelse(is.na(x) | x == "NaN", "-", as.character(x)))

  # Display the table with column alignment and fixed column names
  kable(summary_df, caption = "Data Summary", align = rep('c', ncol(summary_df))) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                  full_width = TRUE) %>%
    scroll_box(width = "100%", height = "400px")
} else {
  cat("No data available for summary statistics.")
}
```

### Outlier Detection

```{r}
if(params$include_stats && length(params$numeric_vars) > 0 && !is.null(params$data)) {
  # Simple outlier detection without complex plots
  cat("### Outlier Analysis Results\n\n")
  
  # Process up to 3 numeric variables
  vars_to_analyze <- params$numeric_vars[1:min(3, length(params$numeric_vars))]
  outliers_found <- FALSE
  
  for(var in vars_to_analyze) {
    if(var %in% names(params$data) && is.numeric(params$data[[var]])) {
      # Get non-missing values
      values <- params$data[[var]][!is.na(params$data[[var]])]
      
      if(length(values) >= 5) {
        # Calculate outlier bounds using the IQR method
        q1 <- quantile(values, 0.25, na.rm = TRUE)
        q3 <- quantile(values, 0.75, na.rm = TRUE)
        iqr <- q3 - q1
        lower_bound <- q1 - 1.5 * iqr
        upper_bound <- q3 + 1.5 * iqr
        
        # Count outliers
        low_outliers <- sum(values < lower_bound)
        high_outliers <- sum(values > upper_bound)
        total_outliers <- low_outliers + high_outliers
        
        # Basic statistics
        min_val <- min(values)
        max_val <- max(values)
        median_val <- median(values)
        
        # Display results
        cat("**", var, "**\n\n", sep="")
        
        if(total_outliers > 0) {
          outliers_found <- TRUE
          cat("* **Outliers detected:** ", total_outliers, " (", 
              round(100 * total_outliers / length(values), 2), "% of values)\n", sep="")
          
          if(low_outliers > 0) {
            cat("* **Low outliers:** ", low_outliers, " values below ", round(lower_bound, 2), "\n", sep="")
          }
          
          if(high_outliers > 0) {
            cat("* **High outliers:** ", high_outliers, " values above ", round(upper_bound, 2), "\n", sep="")
          }
        } else {
          cat("* No outliers detected\n")
        }
        
        cat("* **Range:** ", round(min_val, 2), " to ", round(max_val, 2), "\n", sep="")
        cat("* **Normal range:** ", round(lower_bound, 2), " to ", round(upper_bound, 2), " (based on IQR method)\n", sep="")
        cat("* **Median:** ", round(median_val, 2), "\n\n", sep="")
      } else {
        cat("**", var, "**: Insufficient data for outlier analysis (requires at least 5 non-missing values)\n\n", sep="")
      }
    }
  }
  
  if(!outliers_found && length(vars_to_analyze) > 0) {
    cat("No outliers were detected in the analyzed variables.")
  }
} else {
  cat("Insufficient numeric data for outlier detection.")
}
```
```

## Data Visualizations

### Numeric Variables

```{r}
if(params$include_plots && length(params$numeric_vars) > 0 && !is.null(params$data)) {
  # Histogram for numeric variables (show up to 4)
  num_to_show <- min(4, length(params$numeric_vars))
  for(i in 1:num_to_show) {
    var <- params$numeric_vars[i]
    if(var %in% names(params$data)) {
      tryCatch({
        p <- ggplot(params$data, aes_string(x = var)) +
          geom_histogram(fill = "steelblue", bins = 30) +
          theme_minimal() +
          labs(title = paste("Distribution of", var),
               x = var,
               y = "Frequency")
        print(p)
      }, error = function(e) {
        cat("Could not generate histogram for", var, ":", e$message, "\n")
      })
    }
  }
} else {
  cat("No numerical variables available for visualization.")
}
```

### Categorical Variables

```{r}
if(params$include_plots && length(params$categorical_vars) > 0 && !is.null(params$data)) {
  # Bar plots for categorical (show up to 4)
  cat_to_show <- min(4, length(params$categorical_vars))
  
  for(i in 1:cat_to_show) {
    var <- params$categorical_vars[i]
    if(var %in% names(params$data)) {
      tryCatch({
        # Count unique values to determine if we need to rotate labels
        unique_count <- length(unique(params$data[[var]]))
        angle <- if(unique_count > 5) 45 else 0
        
        p <- ggplot(params$data, aes_string(x = var)) +
          geom_bar(fill = "steelblue") +
          theme_minimal() +
          theme(axis.text.x = element_text(angle = angle, hjust = 1)) +
          labs(title = paste("Distribution of", var),
               x = var,
               y = "Count")
        print(p)
      }, error = function(e) {
        cat("Could not generate bar plot for", var, ":", e$message, "\n")
      })
    }
  }
} else {
  cat("No categorical variables available for visualization.")
}
```

### Correlation Analysis

```{r}
if(params$include_stats && params$include_correlation && length(params$numeric_vars) > 1 && !is.null(params$data)) {
  # Create correlation matrix for numeric variables
  tryCatch({
    # Select numeric columns
    numeric_data <- params$data %>% select(all_of(params$numeric_vars))
    
    # Check if we have enough complete cases for correlation
    complete_rows <- complete.cases(numeric_data)
    if(sum(complete_rows) >= 2) {
      # Create correlation plot
      corr_matrix <- cor(numeric_data, use = "pairwise.complete.obs")
      corrplot::corrplot(corr_matrix, 
                        method = "circle", 
                        type = "upper", 
                        tl.col = "black", 
                        tl.cex = 0.7,
                        title = "Correlation Matrix of Numeric Variables",
                        mar = c(0, 0, 1, 0))
    } else {
      cat("Insufficient complete cases for correlation analysis. Please check for missing values in numeric variables.")
    }
  }, error = function(e) {
    cat("Could not generate correlation plot due to an error:", e$message)
  })
} else {
  cat("Insufficient data for correlation analysis.")
}
```

### Bivariate Relationships

```{r}
if(params$include_plots && length(params$numeric_vars) >= 2 && !is.null(params$data)) {
  # Show scatterplot matrix for first few numeric variables
  tryCatch({
    # Select up to 3 numeric variables to avoid overloading the report
    vars_to_plot <- params$numeric_vars[1:min(3, length(params$numeric_vars))]
    pairs_data <- params$data %>% select(all_of(vars_to_plot))
    
    # Create scatter plot matrix
    GGally::ggpairs(pairs_data, 
                   title = "Relationships Between Numeric Variables")
  }, error = function(e) {
    # If ggpairs fails, try simple scatter plots
    tryCatch({
      if(length(params$numeric_vars) >= 2) {
        # Create a basic scatter plot of the first two numeric variables
        p <- ggplot(params$data, aes_string(x = params$numeric_vars[1], y = params$numeric_vars[2])) +
          geom_point(alpha = 0.5) +
          theme_minimal() +
          labs(title = paste("Relationship between", params$numeric_vars[1], "and", params$numeric_vars[2]))
        print(p)
      }
    }, error = function(e2) {
      cat("Could not generate bivariate plots due to an error:", e2$message)
    })
  })
} else {
  cat("Insufficient numeric variables for bivariate analysis.")
}
```

## Executive Summary

```{r}
if(!is.null(params$data) && nrow(params$data) > 0) {
  # Create an executive summary with key metrics
  cat("Key Dataset Metrics\n\n")
  
  # Data size
  data_size <- format(object.size(params$data), units = "auto")
  num_rows <- nrow(params$data)
  num_cols <- ncol(params$data)
  
  # Missing data percentage - including empty strings and NA text for character columns
  missing_cells <- sum(sapply(params$data, function(x) {
    if(is.factor(x) || is.character(x)) {
      sum(is.na(x) | x == "" | x == "NA", na.rm = TRUE)
    } else {
      sum(is.na(x), na.rm = TRUE)
    }
  }))
  total_cells <- num_rows * num_cols
  missing_pct <- round((missing_cells/total_cells) * 100, 2)
  
  # Variable types
  num_numeric <- length(params$numeric_vars)
  num_categorical <- length(params$categorical_vars)
  
  # Create a simple executive summary table
  summary_table <- data.frame(
    Metric = c("Rows", "Columns", "Data Size", "Missing Data", 
               "Numeric Variables", "Categorical Variables"),
    Value = c(num_rows, num_cols, data_size, paste0(missing_pct, "%"),
              num_numeric, num_categorical)
  )
  
  kable(summary_table, caption = "Dataset Overview") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
} else {
  cat("No data available for executive summary.")
}
```