---
title: "📊DataPrepHIS Report"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: united
    code_folding: hide
params:
  data: NULL
  original_filename: NULL
  preprocessing_summary: NULL
  numeric_vars: NULL
  categorical_vars: NULL
  visualizations: NULL
  include_plots: TRUE
  include_stats: TRUE
  include_preprocessing: TRUE
  include_correlation: TRUE
  include_quality_report: TRUE
  include_recommendations: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dlookr)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(corrplot)
library(GGally)
library(moments)
```

# Data Pre-Processing and Visualization Report 

This report summarizes the data preprocessing steps and visualizations for the tool 📊DataPrepHIS.

## 1. Dataset Overview

```{r}
# Display dataset name
if(!is.null(params$original_filename)) {
  cat("**Dataset Name:** ", params$original_filename)
} else {
  cat("**Dataset Name:** Not specified")
}
```

### 1.1 Data Preview

```{r}
# Basic summary of dataset
if(!is.null(params$data) && nrow(params$data) > 0) {
  kable(head(params$data, 10), 
       caption = "First 10 rows of dataset") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                 full_width = TRUE) %>%
    scroll_box(width = "100%", height = "300px")
} else {
  cat("No data available to display.")
}
```

### 1.2 Dataset Structure

```{r}
# Display data structure
if(params$include_stats && !is.null(params$data) && ncol(params$data) > 0) {
  # Dataset dimensions
  cat("**Dataset Dimensions:** ", nrow(params$data), " rows × ", ncol(params$data), " columns\n\n")
  
  # Variable types table
  params$data %>% 
    diagnose() %>%
    kable(caption = "Data Type Distribution") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
} else {
  cat("No data available for structure analysis.")
}
```

## 2. Data Preprocessing

### 2.1 Preprocessing Operations

```{r}
# Display preprocessing summary table
if(params$include_preprocessing && !is.null(params$preprocessing_summary) && nrow(params$preprocessing_summary) > 0) {
  kable(params$preprocessing_summary, 
       caption = "Applied Preprocessing Operations") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
} else {
  cat("No preprocessing operations were recorded.")
}
```

### 2.2 Missing Values Analysis

```{r}
# Display missing values analysis
if(params$include_stats && !is.null(params$data)) {
  if(sum(is.na(params$data)) > 0) {
    tryCatch({
      params$data %>% 
        plot_na_pareto(main = "Missing Values Distribution")
    }, error = function(e) {
      cat("**Missing value analysis:** There are", sum(is.na(params$data)), "missing values in the dataset.")
    })
  } else {
    cat("**No missing values** found in the dataset.")
  }
}
```

## 3. Data Exploration

### 3.1 Summary Statistics

```{r}
if(params$include_stats && !is.null(params$data) && ncol(params$data) > 0) {
  # Generate summary statistics
  summary_stats <- data.frame(
    Column = names(params$data),
    Class = sapply(params$data, function(x) { if (is.numeric(x)) "Numerical" else "Categorical" }),
    Missing = sapply(params$data, function(x) {
      if (is.factor(x) || is.character(x)) {
        sum(is.na(x) | x == "" | x == "NA", na.rm = TRUE)
      } else {
        sum(is.na(x), na.rm = TRUE)
      }
    }),
    Missing_Percent = sapply(params$data, function(x) {
      missing_count <- if (is.factor(x) || is.character(x)) {
        sum(is.na(x) | x == "" | x == "NA", na.rm = TRUE)
      } else {
        sum(is.na(x), na.rm = TRUE)
      }
      sprintf("%.1f%%", (missing_count / length(x)) * 100)
    })
  )
  
  # Add numeric-specific stats
  summary_stats$Min <- sapply(params$data, function(x) if (is.numeric(x)) round(min(x, na.rm = TRUE), 2) else NA)
  summary_stats$Max <- sapply(params$data, function(x) if (is.numeric(x)) round(max(x, na.rm = TRUE), 2) else NA)
  summary_stats$Mean <- sapply(params$data, function(x) if (is.numeric(x)) round(mean(x, na.rm = TRUE), 2) else NA)
  summary_stats$SD <- sapply(params$data, function(x) if (is.numeric(x)) round(sd(x, na.rm = TRUE), 2) else NA)
  
  # Replace NA values with dash for display
  summary_stats[is.na(summary_stats)] <- "-"
  
  kable(summary_stats, caption = "Summary Statistics") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                 full_width = TRUE) %>%
    scroll_box(width = "100%", height = "500px")
} else {
  cat("No data available for summary statistics.")
}
```

### 3.2 Outlier Detection

```{r}
if(params$include_stats && params$include_plots && length(params$numeric_vars) > 0 && !is.null(params$data)) {
  # Show outlier plot for numeric variables
  tryCatch({
    params$data %>%
      select(params$numeric_vars[1:min(3, length(params$numeric_vars))]) %>%
      plot_outlier()
  }, error = function(e) {
    cat("Could not generate outlier plot due to an error:", e$message)
  })
}
```

## 4. Data Visualizations

### 4.1 Numeric Variables

```{r}
if(params$include_plots && length(params$numeric_vars) > 0 && !is.null(params$data)) {
  # Histogram for numeric variables (show up to 4)
  num_to_show <- min(4, length(params$numeric_vars))
  for(i in 1:num_to_show) {
    var <- params$numeric_vars[i]
    if(var %in% names(params$data)) {
      tryCatch({
        p <- ggplot(params$data, aes_string(x = var)) +
          geom_histogram(fill = "steelblue", bins = 30) +
          theme_minimal() +
          labs(title = paste("Distribution of", var),
               x = var,
               y = "Frequency")
        print(p)
      }, error = function(e) {
        cat("Could not generate histogram for", var, ":", e$message, "\n")
      })
    }
  }
} else {
  cat("No numerical variables available for visualization.")
}
```

### 4.2 Categorical Variables

```{r}
if(params$include_plots && length(params$categorical_vars) > 0 && !is.null(params$data)) {
  # Bar plots for categorical (show up to 4)
  cat_to_show <- min(4, length(params$categorical_vars))
  for(i in 1:cat_to_show) {
    var <- params$categorical_vars[i]
    if(var %in% names(params$data)) {
      tryCatch({
        # Count unique values to determine if we need to rotate labels
        unique_count <- length(unique(params$data[[var]]))
        angle <- if(unique_count > 5) 45 else 0
        
        p <- ggplot(params$data, aes_string(x = var)) +
          geom_bar(fill = "steelblue") +
          theme_minimal() +
          theme(axis.text.x = element_text(angle = angle, hjust = 1)) +
          labs(title = paste("Distribution of", var),
               x = var,
               y = "Count")
        print(p)
      }, error = function(e) {
        cat("Could not generate bar plot for", var, ":", e$message, "\n")
      })
    }
  }
} else {
  cat("No categorical variables available for visualization.")
}
```

### 4.3 Correlation Analysis

```{r}
if(params$include_stats && params$include_correlation && length(params$numeric_vars) > 1 && !is.null(params$data)) {
  # Create correlation matrix for numeric variables
  tryCatch({
    # Select numeric columns
    numeric_data <- params$data %>% select(all_of(params$numeric_vars))
    
    # Create correlation plot
    corr_matrix <- cor(numeric_data, use = "pairwise.complete.obs")
    corrplot::corrplot(corr_matrix, 
                      method = "circle", 
                      type = "upper", 
                      tl.col = "black", 
                      tl.cex = 0.7,
                      title = "Correlation Matrix of Numeric Variables",
                      mar = c(0, 0, 1, 0))
  }, error = function(e) {
    cat("Could not generate correlation plot due to an error:", e$message)
  })
} else {
  cat("Insufficient data for correlation analysis.")
}
```

### 4.4 Bivariate Relationships

```{r}
if(params$include_plots && length(params$numeric_vars) >= 2 && !is.null(params$data)) {
  # Show scatterplot matrix for first few numeric variables
  tryCatch({
    vars_to_plot <- params$numeric_vars[1:min(3, length(params$numeric_vars))]
    pairs_data <- params$data %>% select(all_of(vars_to_plot))
    
    GGally::ggpairs(pairs_data, 
                   title = "Relationships Between Numeric Variables")
  }, error = function(e) {
    cat("Could not generate bivariate plots due to an error:", e$message)
  })
} else {
  cat("Insufficient numeric variables for bivariate analysis.")
}
```

## 6. Executive Summary

```{r}
if(!is.null(params$data)) {
  # Create an executive summary with key metrics
  cat("Key Dataset Metrics\n\n")
  
  # Data size
  data_size <- format(object.size(params$data), units = "auto")
  num_rows <- nrow(params$data)
  num_cols <- ncol(params$data)
  
  # Missing data percentage
  missing_cells <- sum(is.na(params$data))
  total_cells <- num_rows * num_cols
  missing_pct <- round((missing_cells/total_cells) * 100, 2)
  
  # Variable types
  num_numeric <- length(params$numeric_vars)
  num_categorical <- length(params$categorical_vars)
  
  # Create a simple executive summary table
  summary_table <- data.frame(
    Metric = c("Rows", "Columns", "Data Size", "Missing Data", 
               "Numeric Variables", "Categorical Variables"),
    Value = c(num_rows, num_cols, data_size, paste0(missing_pct, "%"),
              num_numeric, num_categorical)
  )
  
  kable(summary_table, caption = "Dataset Overview") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```
